---
title: "ImportantStudies_Plots"
author: "Aïsha Schuhegger"
date: "5/16/2024"
output: html_document
---

This script provides all plots that are used in the paper "Impact of progress visualization and user feedback on the long-term usage of mental health apps".

# Import libraries & set working directory

```{r, comment=NA}
library(dplyr)
library(ggplot2)
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
folderpath = "../figures/"
if (!dir.exists(folderpath)) {
  # If the folder does not exist, create it
  dir.create(folderpath, recursive = TRUE)
  print(paste("Folder", folderpath, "created."))
} else {
  print(paste("Folder", folderpath, "already exists."))
}
```

# Read in CSV File and check that the names of the authors are represented in the rownames

```{r}
studies <- read.csv("../ImportantStudies.csv", sep=";")
rownames(studies) <- studies$Author
colnames(studies)[colnames(studies) == "Length.weeks."] <- "Length"
# Make columns numeric that need to be
studies$RetentionRate <- as.numeric(gsub(",", ".", studies$RetentionRate))
studies$UserSatisfaction <- as.numeric(gsub(",", ".", studies$UserSatisfaction))
studies$ProgressVisualization <- as.numeric(studies$ProgressVisualization)
studies$Feedback <- as.numeric(studies$Feedback)
studies$Length <- as.numeric(studies$Length)

custom_colors <- c("0" = "lightblue2", "1" = "paleturquoise4", "2" = "deepskyblue4")
custom_labels <- c("0" = "Progress \nVisualization", "1" = "User Feedback", "2"  = "Progress Visualization \nAnd User Feedback")
```

# Retention Rate
```{r}

studies$ProgFeedBoth <- as.factor(studies$ProgFeedBoth)

retentionrate_authors <- ggplot(studies, aes(x = Author, y = RetentionRate, fill = ProgFeedBoth)) +
  geom_bar(stat = "identity", width=0.5) +
  ylim(0, 1) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ggtitle("Retention Rates by Author") +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_fill_manual(values = custom_colors, labels = custom_labels) +
  labs(fill = "Type of Factor")  # Adding legend title
retentionrate_authors

ggsave(paste(folderpath, "/retentionrate_authors.png", sep=""), plot = retentionrate_authors, width = 5.5, height = 4.5, units = "in", dpi = 300)
```

# Retention Rate vs. Length
```{r}
retrate_length_scatterplot <- ggplot(studies, aes(x = Length, y = RetentionRate)) +
  geom_point(color="darkblue") +
  labs(title = "Retention Rate vs. Length",
       x = "Length (in weeks)",
       y = "Retention Rate") +
  theme(plot.title = element_text(hjust = 0.5)) 
ggsave(paste(folderpath, "/retentionrate_length.png", sep=""), plot = retrate_length_scatterplot, width = 5.5, height = 4.5, units = "in", dpi = 300)

retrate_length_trend <- ggplot(studies, aes(x = Length, y = RetentionRate)) +
  geom_point(color="darkblue") +
  geom_smooth(method = "lm", se = FALSE, color="grey") +
  labs(title = "Retention Rate vs. Length",
       x = "Length (in weeks)",
       y = "Retention Rate") +
  theme(plot.title = element_text(hjust = 0.5)) 
retrate_length_trend
ggsave(paste(folderpath, "/retentionrate_length_trend.png", sep=""), plot = retrate_length_trend, width = 5.5, height = 4.5, units = "in", dpi = 300)

studies_adjusted <- studies
studies_adjusted <- studies[-c(6),]
retrate_length_adjusted_loweroutlier <- ggplot(studies_adjusted, aes(x = Length, y = RetentionRate)) +
  geom_point(color="darkblue") +
  geom_smooth(method = "lm", se = FALSE, color="grey") +
  labs(title = "Retention Rate vs. Length (without low outlier)",
       x = "Length (in weeks)",
       y = "Retention Rate") +
  theme(plot.title = element_text(hjust = 0.5)) 
ggsave(paste(folderpath, "/retentionrate_length_withoutLowerOutlier.png", sep=""), plot = retrate_length_adjusted_loweroutlier, width = 5.5, height = 4.5, units = "in", dpi = 300)
retrate_length_adjusted_loweroutlier
```

#Retention Rate vs. User Satisfaction
```{r}
studies_clean <- studies %>% 
  filter(!is.na(UserSatisfaction), !is.na(RetentionRate))
studies$SatisfactionCategory <- cut(studies$UserSatisfaction, 
                                    breaks = c(-Inf, 0, 1, Inf), 
                                    labels = c("low", "medium", "high"), 
                                    right = FALSE)
# Create the boxplot
retrate_us <- ggplot(studies, aes(x = factor(SatisfactionCategory), y = RetentionRate)) +
  geom_boxplot(color = "darkblue") +
  labs(
    title = "Retention Rate by User Satisfaction",
    x = "User Satisfaction",
    y = "Retention Rate"
  ) +
  theme(plot.title = element_text(hjust = 0.5))
retrate_us
ggsave(paste(folderpath, "/retentionrate_user_satisfaction.png", sep=""), plot = retrate_us, width = 5.5, height = 4.5, units = "in", dpi = 300)
```

# User Satisfaction
```{r}
usersat <- ggplot(studies, aes(x = factor(SatisfactionCategory), fill = factor(ProgFeedBoth))) +
  geom_bar(width = 0.5) +
  scale_fill_manual(values = custom_colors, labels = custom_labels) +
  labs(title = "User Satisfaction Distribution",
       x = "User Satisfaction",
       y = "Amount of Studies",
       fill = "Type of Factor") +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_y_continuous(breaks = seq(0, max(table(studies$SatisfactionCategory)), by = 1))

usersat
ggsave(paste(folderpath, "/userSatisfaction_factors.png", sep=""), plot = usersat, width = 6, height = 4.5, units = "in", dpi = 300)
```

# Factor Distribution

```{r}
factors_bar <- ggplot(studies, aes(x = factor(ProgFeedBoth), fill = factor(ProgFeedBoth))) +
  geom_bar(width = 0.5) +
  scale_x_discrete(labels = custom_labels) + 
  scale_fill_manual(values = custom_colors, labels = custom_labels) +  # Use custom colors and labels
  labs(title = "Factor Distribution",
       x = "App Includes Factor",
       y = "Amount of Studies",
       fill = NULL) +  # Remove the legend title
  theme(plot.title = element_text(hjust = 0.5),
        legend.position = "none") +
  scale_y_continuous(breaks = seq(0, 10, by = 1), limits = c(0, 10))  # Set y-axis breaks and limits

factors_bar
ggsave(paste(folderpath, "/AmountStudies_Factors.png", sep=""), plot = factors_bar, width = 5.5, height = 4.5, units = "in", dpi = 300)
```

# Retention Rate vs. grouped factors

```{r}
studies$ProgFeedBoth <- as.numeric(studies$ProgFeedBoth)
studies$FactorsCategory <- cut(studies$ProgFeedBoth, 
                                breaks = c(0, 1, 2), 
                                labels = c("User Feedback", "Progress Visualization \nAnd User Feedback"), 
                                right = TRUE)


retrate_factors <- ggplot(studies, aes(x = factor(FactorsCategory), y = RetentionRate)) +
  geom_boxplot(color = "darkblue") +
  labs(
    title = "Retention Rate by Factor",
    x = "Factor",
    y = "Retention Rate"
  ) +
  theme(plot.title = element_text(hjust = 0.5))
retrate_factors #0.87, 0.88 = median
ggsave(paste(folderpath, "/retentionrate_factors.png", sep=""), plot = retrate_factors, width = 5.5, height = 4.5, units = "in", dpi = 300)
```

